### 2023-09-14 article/article.qmd
add baseline for country information

---
title: "SDG contributions"
format: html
---

```{r load data}
library(sparklyr)
library(ggplot2)
library(patchwork)
library(dplyr)
library(arrow)
library(here)
library(tidyr)
library(forcats)
library(readr)

theme_set(theme_bw())

source(here("R/helpers.R"))

Sys.setenv(SPARK_HOME = "/home/tklebel/spark-3.4.0-bin-hadoop3/")
Sys.setenv(HADOOP_HOME = "/home/hadoop/hadoop-3.3.1")
Sys.setenv(HADOOP_CONF_DIR = "/home/hadoop/hadoop-3.3.1/etc/hadoop")
Sys.setenv(YARN_HOME = "/home/hadoop/hadoop-3.3.1")
Sys.setenv(YARN_CONF_DIR = "/home/hadoop/hadoop-3.3.1/etc/hadoop")
Sys.setenv(JAVA_HOME="/usr/lib/jvm/java-1.11.0-openjdk-amd64")

config <- spark_config()
config$spark.executor.cores <- 5
config$spark.executor.instances <- 38
config$spark.executor.memory <- "20G"
config$spark.hadoop.mapreduce.fileoutputcommitter.algorithm.version <- 2

sc <- spark_connect(master = "yarn", config = config,
                    app_name = "SDG-knowledge-production")

message("Connection to Spark successful!")

message("Reading the datasets...")

# Overall MAG data
paper_cols <- c("paperid", "rank", "doi", "type", "title_normalised", "title", 
                "unknown1", "year", "date", "unknown2", "publisher", 
                paste0("unknown", 3:12),
                "journal", paste0("unknown", c(14:16)))

mag_2021_papers <- spark_read_csv(sc, "/tklebel/mag-2021-03-15/Papers.txt",
                                  name = "mag2021_papers",
                                  memory = FALSE,
                                  delimiter = "\\t", header = FALSE,
                                  columns = paper_cols)

affil_cols <- c("paperid", "authorid", "affiliationid", "authorsequencenumber",
                "originalauthor", "originalaffiliation")

mag_2021_paper_author_affil <- spark_read_csv(
  sc, 
  "/tklebel/mag-2021-03-15/PaperAuthorAffiliations.txt",
  name = "mag2021_paper_author_affil", delimiter = "\\t",
  memory = FALSE, header = FALSE, columns = affil_cols)

# this is data to fill in country information where we do not have it from the
# SDG set
missing_country_data <- spark_read_csv(
  sc,
  "/tklebel/SDG/additional_country_information.csv",
  name = "missing_country_data"
)

# SDG specific data
papers <- spark_read_csv(sc, "/tklebel/SDG/sdg_papers_collated.csv",
                         name = "papers")
sdg_labels <- spark_read_csv(sc, "/tklebel/SDG/sdg_labels.csv",
                             name = "sdg_labels")
# join the labels here, since we are doing everything by SDG
papers <- papers %>% 
  left_join(sdg_labels)

# remove 2020
papers <- papers %>% 
  filter(year < 2020)

# Affiliations
affils <- spark_read_csv(sc,
                         "/tklebel/SDG/affiliations_with_country_code.csv",
                         name = "affils")

author_paper_affiliations <- spark_read_csv(
  sc,
  "/tklebel/SDG/sdg_author_paper_affil.csv",
  name = "author_paper_affiliations"
)

# UN country information
un_countries <- read_csv2(here::here("data/external/UNSD — Methodology.csv"))

un_countries_selection <- un_countries %>% 
  select(country_code = `ISO-alpha3 Code`, country_name = `Country or Area`, 
         continent = `Region Name`, sub_continent = `Sub-region Name`)

```

# Sample size
```{r}
# total sample size
papers %>% 
  sdf_nrow()
```

```{r}
# per SDG
papers %>% 
  count(SDG_label)
```

```{r}
papers %>% 
  group_by(year) %>% 
  filter(year %in% c(2006, 2018)) %>% 
  count(SDG_label) %>% 
  arrange(SDG_label, year)
```

# Growth of literature
```{r sdg_count, fig.width=7, fig.height=4}
sdg_counts <- papers %>% 
  count(SDG_label, year) %>% 
  collect()

overall_count <- mag_2021_papers %>% 
  filter(year %in% 2006:2019) %>% 
  count(year, name = "overall_n") %>% 
  collect() %>% 
  mutate(year = as.integer(year))

merged_count <- sdg_counts %>% 
  left_join(overall_count) %>% 
  mutate(prop = n / overall_n)

merged_count %>% 
  drop_na() %>% 
  ggplot(aes(as_year(year), prop, colour = fix_sdg(SDG_label))) +
  geom_line() +
  geom_point() +
  colorspace::scale_color_discrete_qualitative() +
  scale_y_log10(labels = scales::comma,
                breaks = c(10e+3, 30e+3, 1e+5, 1e+6)) +
  labs(x = NULL, y = "# of publications", colour = NULL,
       title = "Development of SDG areas over time") 
```

Normalised by the overall number of papers in MAG, there is some growth for 
these literatures.

```{r growth-rates}
growth_rate_data <- merged_count %>% 
  arrange(SDG_label, year) %>% 
  group_by(SDG_label) %>% 
  mutate(prop_change = prop / lag(prop),
         prop_change = replace_na(prop_change, 1),
         prop_change = prop_change - 1)

growth_rate_data %>% 
  ggplot(aes(as_year(year), prop_change, colour = fix_sdg(SDG_label))) +
  geom_line() +
  geom_point() +
  colorspace::scale_color_discrete_qualitative() +
  scale_y_continuous(labels = scales::percent) +
  labs(x = NULL, y = "Yearly growth rate relative to total papers in MAG",
       colour = NULL,
       title = "Development of SDG areas over time") +
  theme(legend.position = "top")
```

```{r}
growth_rate_data %>% 
  knitr::kable()

growth_rate_data %>% 
  filter(year %in% c(2006, 2019)) %>% 
  summarise(overall_growth_factor = {overall_n / lag(overall_n)}[2],
            individual_growth_factor = {n / lag(n)}[2],
            individual_growth_rate = individual_growth_factor^(1/(2019-2006)) - 1,
            overall_growth_rate = overall_growth_factor^(1/(2019-2006)) - 1) %>% 
  knitr::kable()

```

# Continents over time
```{r}
author_paper_affiliations_w_groups <- make_author_groups(author_paper_affiliations)

papers_w_affils <- papers %>% 
  select(paperid, SDG_label, year, citations_norm) %>% 
  left_join(author_paper_affiliations_w_groups) %>% 
  left_join(affils) %>% 
  select(paperid, SDG_label, year, authorid, affiliationid, country,
         paper_author_cat, author_position, citations_norm)
```

```{r}
papers_per_country_fos_author_pos <- papers_w_affils %>% 
  group_by(country, year, SDG_label, author_position) %>% 
  summarise(n = n()) %>% 
  collect()

papers_per_country_fos_author_pos_country <- papers_per_country_fos_author_pos %>%
  left_join(un_countries_selection, by = c("country" = "country_code")) %>% 
  drop_na()

normalised_region_share_time <- papers_per_country_fos_author_pos_country %>% 
  group_by(SDG_label, year, sub_continent) %>% 
  summarise(nn = sum(n)) %>% 
  mutate(share = nn/sum(nn))
```

```{r sub_continents_over_time, fig.width=8.2, fig.height=4.4}
date_scale <- scale_x_date(breaks = as_year(c(2006, 2010, 2015, 2019)),
               date_labels = "%Y")
base_col = "grey80"

subset_countries <- filter(
  normalised_region_share_time, 
  sub_continent %in% c("Northern America", "Western Europe", "Northern Europe",
                       "Eastern Asia"
  ))

normalised_region_share_time %>% 
  ggplot(aes(as_year(year), share, group = sub_continent, colour = sub_continent)) +
  geom_line(colour = base_col) +
  geom_line(data = subset_countries) +
  geom_point(data = subset_countries, size = 1.2) +
  facet_wrap(vars(fix_sdg(SDG_label))) +
  date_scale +
  scale_y_continuous(labels = function(x) scales::percent(x, 1)) +
  colorspace::scale_color_discrete_qualitative() +
  theme(legend.position = "top") +
  labs(x = NULL, y = "% of publications from world region within SDGs",
       colour = NULL)
```

This is now equivalent to full counting, since all author positions are included.

## Baseline calculation

```{r}
mag_2021_papers_w_affils <- mag_2021_papers %>% 
  select(paperid, year) %>% 
  left_join(mag_2021_paper_author_affil) %>% 
  left_join(affils) %>% 
  left_join(missing_country_data) %>% 
  select(paperid, year, authorid, affiliationid, country)
```

```{r}
mag_2021_papers_per_country <- mag_2021_papers_w_affils %>% 
  filter(year %in% 2006:2019) %>% 
  group_by(country, year) %>% 
  summarise(n = n()) %>% 
  collect()

mag_2021_papers_per_country_w_country_info <- mag_2021_papers_per_country %>%
  left_join(un_countries_selection, by = c("country" = "country_code")) %>% 
  drop_na()

mag_2021_normalised_region_share_time <- mag_2021_papers_per_country_w_country_info %>% 
  group_by(year, sub_continent) %>% 
  summarise(nn = sum(n)) %>% 
  mutate(share = nn/sum(nn))
```

```{r baseline_sub_continents_over_time, fig.width=8.2, fig.height=4.4}
date_scale <- scale_x_date(breaks = as_year(c(2006, 2010, 2015, 2019)),
               date_labels = "%Y")
base_col = "grey80"

subset_countries <- filter(
  mag_2021_normalised_region_share_time, 
  sub_continent %in% c("Northern America", "Western Europe", "Northern Europe",
                       "Eastern Asia"
  ))

mag_2021_normalised_region_share_time %>% 
  ggplot(aes(as_year(year), share, group = sub_continent, colour = sub_continent)) +
  geom_line(colour = base_col) +
  geom_line(data = subset_countries) +
  geom_point(data = subset_countries, size = 1.2) +
  date_scale +
  scale_y_continuous(labels = function(x) scales::percent(x, 1)) +
  colorspace::scale_color_discrete_qualitative() +
  theme(legend.position = "top") +
  labs(x = NULL, y = "% of publications from world region across MAG",
       colour = NULL)
```

The above is the same graph above but for all of MAG. The next step is to 
integrate this into the existing graph. Options:

- Normalise existing graph by baseline
- Add baseline as a background (maybe with different line type)

The latter seems more in line with the focus of the analysis, since the target
is the overall distribution of knowledge production, not the relative 
specialisation.

```{r}
spark_disconnect(sc)
```